---
title: "Time series"
author: "Julian Flowers"
date: "05/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NHSRdatasets); library(tidyverse);library(lubridate);library(mgcv);library(rstanarm); library(lme4)
```

## 

```{r}
ae <- NHSRdatasets::ae_attendances
ae <- ae %>%
  drop_na() %>%
  filter(type == 1)

mod1 <- glm(breaches ~ as.numeric(period) + admissions + offset(attendances), data = ae, family = poisson)


```





range(ae$period)

length(unique(ae$org_code))

glimpse(ae)

## just look at type 1 AEs

ae_type_1 <- ae %>%
  filter(type == "1")

## summarise

ae_type_1_summary <- ae_type_1 %>%
  group_by(period) %>%
  summarise(tot_att = sum(attendances), 
            tot_breach = sum(breaches), 
            tot_admissions = sum(admissions)) %>%
  mutate(conv_ratio = tot_admissions / tot_att, 
         breach_ratio = tot_breach / tot_att)


ae_type_1_summary %>%
  ggplot() +
  geom_smooth(aes(period, conv_ratio), method = "gam") + 
  geom_point(aes(period, conv_ratio)) +
  geom_smooth(aes(period, breach_ratio), method = "gam")  ## note that is calculates correct aspect ratio and banks data at 45 degrees after Cleveland.

ae_type_1_summary %>%
  ggplot(aes(conv_ratio, breach_ratio)) +
  geom_point() +
  geom_smooth(method = "lm")
  
  
  
  

```

```{r}

ae_type_1_prov <- ae_type_1 %>%
  mutate(conv_ratio = admissions / attendances, 
         breach_ratio = breaches / attendances)

library(tsibble)

ae_type_1_prov_ts <- tsibble::as_tsibble(ae_type_1_prov, key = org_code, index = period)
```

```{r}

model <- bam(breach_ratio ~ s(conv_ratio) + period, data = ae_type_1_summary)

broom.mixed::tidy(model) %>%
  mutate(id = row_number()) %>%
  filter(p.value < 0.01)

```


## LMER

```{r}

str(ae_type_1_prov)

mod_1 <- lmer(breach_ratio ~ 1 + (1|org_code), 
              data = ae_type_1_prov, 
              REML = FALSE)

mod_1


```

```{r}

mod_2 <- lmer(breach_ratio ~ 1 + conv_ratio + (1|org_code), 
              data = ae_type_1_prov, 
              REML = FALSE)

mod_2


```

```{r}

mod_3 <- lmer(breach_ratio ~ 1 + conv_ratio + (1 + conv_ratio|org_code), 
              data = ae_type_1_prov, 
              REML = FALSE)

mod_3


```


```{r}

J <- length(unique(ae_type_1_prov$org_code))
N <- nrow(ae_type_1_prov)

## complete pooling

pooled <- lm(breach_ratio ~ conv_ratio, data = ae_type_1_prov)
a_pooled <- coef(pooled)[1]
b_pooled <- coef(pooled)[2]

## no pooling

nopooled <- lm(formula = breach_ratio ~ 0 + org_code + conv_ratio,
               data = ae_type_1_prov)
a_nopooled <- coef(nopooled)[1:J]   # 73 no-pooling intercepts              
b_nopooled <- coef(nopooled)[J+1]

# Partial pooling (multilevel) regression
a_part_pooled <- coef(mod_3)$org_code[, 1]
b_part_pooled <- coef(mod_3)$org_code[, 2]



```

```{r}

y <- ae_type_1_prov$breach_ratio
x <- as.numeric(ae_type_1_prov$conv_ratio) 
org_id <- ae_type_1_prov$org_code

set.seed(1234)

org_sel <- sample(org_id, 10)

df <- data.frame(y, x, org_id)
df8 <- subset(df, org_id %in% org_sel)

df8$a_pooled <- a_pooled 
df8$b_pooled <- b_pooled
df8$a_nopooled <- a_nopooled[df8$org_id]
df8$b_nopooled <- b_nopooled
df8$a_part_pooled <- a_part_pooled[df8$org_id]
df8$b_part_pooled <- b_part_pooled[df8$org_id]

ggplot(data = df8, 
       aes(x = x, y = y)) + 
  facet_wrap(facets = ~ org_id, 
             ncol = 5) + 
  theme_bw() +
  geom_jitter(position = position_jitter(width = .05, 
                                         height = 0)) +
  geom_abline(aes(intercept = a_pooled, 
                  slope = b_pooled), 
              linetype = "solid", 
              color = "blue", 
              size = 0.5) +
  geom_abline(aes(intercept = a_nopooled, 
                  slope = b_nopooled), 
              linetype = "longdash", 
              color = "red", 
              size = 0.5) + 
  geom_abline(aes(intercept = a_part_pooled, 
                  slope = b_part_pooled), 
              linetype = "dotted", 
              color = "purple", 
              size = 0.7) 



+ 
  scale_x_continuous(breaks = c(0, 1), 
                     labels = c("male", "female")) + 
  labs(title = "Complete-pooling, No-pooling, and Partial pooling estimates",
       x = "", 
       y = "Total score on coursework paper")+theme_bw( base_family = "serif")


```



```{r}

data("roaches")

mod_a <- rstanarm::stan_glmer(breach_ratio ~ 1 + (1|org_code), data = ae_type_1_prov, cores = 6)

prior_summary(mod_a)

sd(ae_type_1_prov$breach_ratio)

print(mod_a)

summary(mod_a, 
        pars = c("(Intercept)", "sigma", "Sigma[org_code:(Intercept),(Intercept)]"),
        probs = c(0.025, 0.975),
        digits = 2)


sims <- as.matrix(mod_a)
cnames <- colnames(sims)

mu_a_sims <- as.matrix(mod_a, 
                       pars = "(Intercept)")
# draws for 73 schools' school-level error
u_sims <- as.matrix(mod_a, 
                    regex_pars = "b\\[\\(Intercept\\) org_code\\:")
# draws for 73 schools' varying intercepts               
a_sims <- as.numeric(mu_a_sims) + u_sims          

# Obtain sigma_y and sigma_alpha^2
# draws for sigma_y
s_y_sims <- as.matrix(mod_a, 
                       pars = "sigma")
# draws for sigma_alpha^2
s__alpha_sims <- as.matrix(mod_a, 
                       pars = "Sigma[org_code:(Intercept),(Intercept)]")

a_mean <- map_dbl(a_sims, mean)
a_sd <- map_dbl(a_sims, sd, na.rm = TRUE)
a_q <- map(a_sims, ~(quantile(.x, na.rm = TRUE, probs = c(.025, .5, .975))))

mod_b <- rstanarm::stan_glmer(breach_ratio ~ conv_ratio + 1|org_code, data = ae_type_1_prov, cores = 6)

summary(mod_b)

broom::tidy(mod_b, intervals = TRUE, parameters = "hierarchical")

  prior_summary(mod_b)
pp_check(mod_b)

launch_shinystan(mod_b)

ppd <- posterior_predict(mod_b, newdata = ae_type_1_prov)

ppd_df <- data.frame(ppd, each = nrow(ppd))


```




