---
title: "aRt of the possible"
subtitle: "Using R"
author: "Julian Flowers"
date: "13th Jan 2022 (updated: `r Sys.Date()`)"
output:
  html_document:
    toc: yes
    code_folding: hide
    
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE)
library(pacman)
p_load(tidyverse)


```


## Why learn R?

* Transformational
* Extend what you can do
* Marketable
* Added value
* Huge community of support
* Its the future and the future is now
* Its fun!



## Environments and systems 
### R is:
* A statistical progamming language
* Open source
* An essential toolkit for data science and modern analysis
    * A GIS
    * A modelling toolkit
    * A machine learning toolkit
    * A time series analysis toolkit
    * A suite of data wrangling tools
    * A suite of high quality visualisation tools
    * A set of tools for creating interactive vizzes and dashboards
    * A powerful text mining toolkit
    * A set of tools for undertaking reproducible open science
    * A tool for creating and building analytical and data services
    

## Getting up and running (quickly)

* Download and install:
    + [R (latest version 4.1.2)](https://cran.r-project.org/src/base/R-4/)
    + [RTools (v4) if running on Windows](https://cran.r-project.org/bin/windows/Rtools/rtools40.html)
    + [RStudio IDE](https://www.rstudio.com/products/rstudio/download/#download) (latest is RStudio Desktop 2022.02.1+461)
    
* Set up a Github account
      + Version control system
      + This can be organisational or personal; public or private
      
* Learn how to:
    + Install packages
          + CRAN
          + Ropensci
          + From Github
    + Get data into R
    + Use `dplyr`, `ggplot2`
    + Write simple functions
    + Use RMarkdown
  

### Getting data into R

* Can import data is any format e.g.
  + csv 
     - `readr::read_csv()`
     - `data.table::fread()` -  fast read for large files
  + xlsx 
     - `readxl::read_excel()`
  + ods
     - `ODS::read_ods()`
  + .dta / .sav
     - `haven::read_sav()` / `haven::read_spss()`
     - `haven::read_dta()`
  + text files
     - readr::read_delim()
     
* R also has its own data format `.rds` which is fast and highly compressed

```{r}


data <- read_csv("https://coronavirus.data.gov.uk/api/v1/data?filters=areaType=nation;areaName=England&structure=%7B%22areaType%22:%22areaType%22,%22areaName%22:%22areaName%22,%22areaCode%22:%22areaCode%22,%22date%22:%22date%22,%22uniqueCasePositivityBySpecimenDateRollingSum%22:%22uniqueCasePositivityBySpecimenDateRollingSum%22,%22uniquePeopleTestedBySpecimenDateRollingSum%22:%22uniquePeopleTestedBySpecimenDateRollingSum%22%7D&format=csv", show_col_types = F)


```

```{r}

head(data)

```
## Core tidy functions {.tabset .tabset-pills}

### `dplyr`

Data wrangling package with a set of key functions (verbs)

### `filter` - operations to select rows on the basis of specified criteria

```{r}

data %>%
  filter(date == "2022-01-01")

```


### `select` -  select columns

```{r}

data %>%
  select(date, uniqueCasePositivityBySpecimenDateRollingSum) %>%
  filter(date >= "2022-01-01")

```

### `mutate` - add a new column

```{r}

data %>%
  mutate(cases = uniqueCasePositivityBySpecimenDateRollingSum / uniqueCasePositivityBySpecimenDateRollingSum[1]) %>%
  select(-areaType) %>%
  head()



```

### Plotting with `ggplot2`

```{r fig.height=4, fig.width=6}

data %>%
  mutate(cases = uniqueCasePositivityBySpecimenDateRollingSum / uniqueCasePositivityBySpecimenDateRollingSum[1]) %>%
  ggplot(aes(date, cases)) +
  geom_line(lty = "dotted") +
  geom_smooth(method = "gam", se = FALSE) +
  theme_bw()



```


### NE Data

```{r, echo = TRUE}

devtools::install_github("ropensci/fingertipsR")

library(fingertipsR); library(tidyverse)

#fingertipsR::select_indicators()

smoking <- fingertips_data(IndicatorID = c(92443, 93085, 93553, 93579, 
                                           93573, 93798, 91547, 92304), 
                           AreaTypeID = 401)


smoking_ne <- smoking %>%
  filter(ParentCode == "E12000001")


```


### Smoking prevalence trend in NE districts

```{r, echo = FALSE}

smoking_ne %>%
  filter(str_detect(IndicatorName, "[Pp]revalence"), Age == "18+ yrs", Sex == "Persons") %>%
  mutate(IndName = str_extract(IndicatorName, "current smokers.*")) %>%
  ggplot(aes(round(TimeperiodSortable/10^4, 0), Value, colour = IndName)) +
  geom_line(aes(group = IndName)) +
  geom_point() +
  geom_linerange(aes(ymin = LowerCI95.0limit, ymax = UpperCI95.0limit)) +
  facet_wrap(~AreaName, 
             nrow = 3) +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 0, hjust =1)) +
  labs(x = "Year", 
       y = "Estimate (%)") +
  scale_x_continuous(n.breaks = 6)



```

---
### Code

```
smoking_ne %>%
  filter(str_detect(IndicatorName, "[Pp]revalence"), Age == "18+ yrs", 
  Sex == "Persons") %>%
  ggplot(aes(round(TimeperiodSortable/10^4, 0), Value, 
  colour = IndicatorName)) +
  geom_line(aes(group = IndicatorName)) +
  geom_point() +
  geom_linerange(aes(ymin = LowerCI95.0limit, ymax = UpperCI95.0limit)) +
  facet_wrap(~AreaName, 
             nrow = 3) +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 0, hjust =1)) +
  labs(x = "Year", 
       y = "Estimate (%)") +
  scale_x_continuous(n.breaks = 6)
```


